### 2019-11-19 cover-letter/cover-letter.tex
Cover letter added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
% It was modified by Katy Huff
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%    PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}

%--------------------------------------------------------------------------------------
%--------INPUT DATA
%--------------------------------------------------------------------------------------
\usepackage{xspace}
\newcommand{\StudentFirstName}{O.\xspace}
\newcommand{\StudentLastName}{Ashraf\xspace}
\newcommand{\RecipientName}{Prof. Mostafa Ghiaasiaan\xspace}
\newcommand{\RecipientAddress}{Executive Editor\\Annals of Nuclear Energy}
%\newcommand{\<++>}{<++>\xspace}

\begin{document}
	\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

	%----------------------------------------------------------------------------------------
	%    ADDRESSEE SECTION
	%----------------------------------------------------------------------------------------

	\begin{letter}{\RecipientName\\
			\RecipientAddress\xspace}

		%-------------------------------------------------------------------------------
		%    YOUR NAME & ADDRESS SECTION
		%-------------------------------------------------------------------------------
		\address{O. Ashraf\\
			oabdelaziz@mephi.ru\\
			osama.ashraf@edu.asu.edu.eg\\
			Institute of Nuc. Physics and Eng.\\
			National Research Nuclear University\\
			Moscow, Russia, 115409}

		%-------------------------------------------------------------------------------
		%    LETTER CONTENT SECTION
		%-------------------------------------------------------------------------------

		\opening{To \RecipientName,}

		Please find enclosed a manuscript entitled: ``Various strategies for thorium fuel cycle transition in the SD-TMSR'' which I and my coauthors (Mr. Andrei Rykhlevskii, Prof. G. V. Tikhomirov, 
		and Prof. Kathryn D. Huff) are submitting for publication as a research article 
		in Annals of Nuclear Energy. The presented manuscript is 100\% original and never been published in any form (symposia, conferences, etc).

		This manuscript investigates the fuel cycle and neutronics performance of the single-fluid double-zone thorium molten salt reactor (SD-TMSR) with different fissile material loadings at startup: High Assay Low Enriched Uranium (HALEU) (19.79\%), Pu mixed with HALEU (19.79\%), reactor-grade Pu, transuranic elements (TRU), and $^{233}$U, using SERPENT-2 Monte Carlo code. This manuscript provides feasible strategies for thorium fuel cycle transition in the SD-TMSR based on reactor-grade Pu and transuranic elements (TRU) results.

		Declarations of interest: none\\

		Thank you for your consideration of our work.

		Your quick response will be highly appreciated.

		\closing{Sincere regards,
			\includegraphics[height=1.5cm]{signature.png}\\
			\fromsig{O. Ashraf\\
				Assistant Teacher\\
				Institute of Nuc. Physics and Eng.\\
				National Research Nuclear University (MEPhI)}
		}

		%-------------------------------------------------------------------------------

	\end{letter}

\end{document}