### 2019-11-25 paper/conclusion.tex

\section{Conclusion} \label{Conclusion}
Five initial fissile loadings strategies have been studied for 
transitioning to the thorium fuel cycle in the Single-fluid Double-zone 
Thorium-based Molten Salt Reactor (SD-TMSR). We 
adopted two different feed mechanisms: thorium and non-thorium. 
This work determined that High Assay Low Enriched Uranium (HALEU), even mixed with Pu, is infeasible. 
Full lifetime (60 year) depletion for the whole-core SD-TMSR model was 
performed with reactor-grade Pu, TRU, and $^{233}$U as initial fissile 
materials. Additionally, the dynamics of the effective multiplication factor 
$k_{eff}$, major isotopes mass, neutron energy spectrum, and essential safety 
parameters have been investigated.

Results demonstrate that continuous flow of reactor-grade Pu allows the 
transition to the thorium fuel cycle in a relatively short time ($\approx$ 
$4.5$ years) compared to $26$ years for Th/$^{233}$U startup fuel. 
Meanwhile, using \gls{TRU} as initial fissile materials shows the possibility 
of operating the SD-TMSR for an extended time ($\approx$ $40$ years) 
without any external feed of $^{233}$U. Notably, the Pu mole concentration (\%) in 
fuel salt was found to be below the solubility limit.

The neutron energy spectrum shift during the reactor operation 
for the Pu and TRU cases is different from the $^{233}$U fueling scenario. 
The spectrum hardens for the $^{233}$U initial fissile isotope during 
operation, but thermalizes for the Pu and TRU cases. Notably, the most 
significant neutron energy spectrum shift was obtained for reactor-grade Pu 
startup loading.

We compared the operational and safety parameters of the \gls{SD-TMSR} for the 
three promising startup fuels at both initial and equilibrium states. The total 
temperature coefficient of reactivity is negative and relatively large in all 
cases. For the TRU case, the coefficient remained almost constant during 
operation: $-4.79\pm0.12$ $pcm/K$ and $-4.76\pm0.11$ $pcm/K$ for the initial 
and equilibrium states, respectively. For reactor-grade Pu, the coefficient 
evolved from $-6.54\pm0.16$ $pcm/K$ to $-4.79\pm0.12$ during 
60 years of operation. Finally, the six factor evolution during the operation 
was calculated for all three cases, providing a basis for designing the reactivity control system of the \gls{SD-TMSR}.

\section{Future work}
The authors intend to verify obtained results using another tool: batch-wise 
code SaltProc \cite{rykhlevskii_arfc/saltproc_2018,rykhlevskii_milestone_2019}.
Additionally, the accident safety analysis using coupled 
neutronics/thermal-hydraulics software, Moltres 
\cite{lindsay_introduction_2018} will be performed. The full-core SERPENT-2 
model of the \gls{SD-TMSR} and equilibrium fuel salt compositions, obtained in 
this work, will be employed to generate problem-oriented nuclear data 
libraries for Moltres. The ultimate goal of this effort is to develop a 
fast-running computational model for studying the dynamic behavior of generic 
\glspl{MSR}, performing safety analysis for different accident scenarios, and 
optimizing the design of various reactor concepts.

\section{Declaration of Competing Interest}

The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.